---
title: "Assignment 4: Data transformation with dplyr and visualization with ggplot "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Instructions: Please read through this before you begin**

* This assignment is due by **10pm on Monday 10/19/20**.

* For this assignment, please **reproduce this markdown file** using R markdown. This includes the followings:

  + **Reproduce this markdown template.** Pay attention to all the formatin gin this file, including bullet points,       bolded characters, inserted code chunks, headings, text colors, blank lines, etc.

  + **Transform the data as instructed.** Try to use `tidyverse` functions even if you are more comfortable with         base-R solutions.Show **the first 6** lines of the transformed data in a table through RMarkdown **using the         kable() function**, as shown in this markdown file.
  
  + **Reproduce the plots exactly as shown in this html file.** In two cases where the plot is not shown (Excercises     3.7 and 3.9), generate plots that you think can best answer the question.
  + Have all your code embedded within the R markdown file, and show **BOTH yourcode and plots** in the knitted           markdown file
  
  + When a verbal response is needed, answer by editing the part in the R markdwon template where it says "Write your    response here".
  
  + Use R Markdwon functionalities to **hide messages and warnings when needed**. (Suggestion: messages and warnings    can often be informative and important, so please examine them carefully and only turn them off when you finish     the exercise).
  
* Please name your R markdown file `assignment_4.Rmd` and the knitted markdown file `assignmrent_4.md`. Plese push     both files to your class GitHub repository.
  
First, load all the required packages with the following code. Install them if they are not installed yet.
  
```{r, message=FALSE}
library(tidyverse)
library(knitr)
library(gapminder) #
```

  
**Exercise 1. Theopylline experiment** 
  
This excercise uses the `Theoph` data frame (comes with your R installation), which has 132 rows and 5 columns of data from an experiment on the pharmacokinetics of the anti-asthmatic drug theophylline. Tweleve subjects were given oral doses of theophylline then serum concentrations were measured at 11 time points over the next 25 hours. You can learn more about this dataset by running `?Theoph` 

Have a look at the data structure

```{r}
kable(head(Theoph))
```

**1.1 Select columns that contain a lower case"t" in the `Theoph` dataset. Do not mannually list all the columns to include.**



```{r}
kable(head(Theoph %>% 
  select(Subject, Wt)))
```




```{r}
kable(head(select(Theoph, contains(('t'))))) # this selects the upper case T why?
```



**Rename the `Wt` column to `Weight` and `conc` column to `Concentration` in the `Theoph` dataset.**

```{r}
kable(head(Theoph %>% 
select(everything(), Weight = Wt, Concentration =conc)))

```


**Extract the `Dose` greater than 4.5 and `Time` greater than the mean `Time`.**

```{r}
kable(head(Theoph %>% 
  filter(Dose > 4.5, Time > mean(Time))))
```



**1.4 Sort the `Theoph` dataset by `Wt` from smallest to largest and secondarily by Time from largest to smallest**

```{r}
kable(head(Theoph %>% 
  arrange(Wt, desc(Time))))

```

  
**1.5 Create a new column called `Quantity`that equals to`Wt` x `Dose` in the `Theoph` dataset. This will tell you the absolute quantity of drug administered to the subject (in mg). Replace the `Dose` Variable**

```{r}
kable(head(Theoph %>% 
  mutate(Quantity=Wt*Dose) %>% 
  select(Subject,Quantity,Time, conc)))
  

```

  
**1.6 Find the mean `conc` and sum of the `Dose` received by each test subject.**

Show data for the 6 subjects with the smallest sum of `Dose` as below ** Do not define new intermediate objects for this excercise; use pipes to chain together functions. **


```{r}
Theoph %>% 
  group_by(Subject) %>% 
  summarize(mean(conc),sum(Dose)) %>% 
  arrange(`sum(Dose)`) %>% 
  head() %>% 
  kable
  
```



**Exercise 2. Trend in land value**

This exercise uses a dataset that describes the trend of land value (`Land.Value`), among other variables in different states in the US 1975-2013.The states are grouped into four different regions, under the variable `regon`. This dataset was obtained from the Data Science Services of Harvard University.


```{r}

housing <- read_csv("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/landdata_states.csv")
kable(head(housing)) 

```


**2.1 Washington DC was not assigned to a region in this dataset. According to the United States Census Bureau, however, DC is part of the South region. Here:**

  * **Change the region of DC to "South" (Hint: there are multiple ways to do this, but `mutate()` and `ifelse()` might be helpful)**
  
  * **Save this updated `region` variable toghether with `state`, `Date` and `Land.Value` into a new data frame**
  
  * **Select the records from DC in this new data frame. How many records are there from DC? Show its first 6 lines.**
  

```{r}

region_updated<-housing %>% 
  mutate(region = ifelse(State == "DC", "South", region)) %>% 
  select(region,State,Land.Value,Date) 
  
  region_updated %>% 
    filter(State=="DC") %>% 
    head() %>% 
    kable() 
    

```
  
  
 
```{r}
DC<- region_updated %>% 
     filter(State=="DC") 
     dim(DC) #or nrow(DC)
```
  
Answer:Write your response here. 
 
The number of records are 153

**2.2 Generate a dataframe that summarizes the mean land value of each region at each time point.**


```{r}

 region_updated %>% 
  group_by(region,Date) %>% 
  summarize (mean_land_value = mean(Land.Value)) %>% 
  head() %>% 
  kable ()

```
 
**2.3 Using the dataframe above, plot the trend in mean land value of each region through time.** 

```{r}
 region_updated %>% 
    group_by(region,Date) %>% 
summarize (mean_land_value = mean(Land.Value)) %>% 
  ggplot(mapping=aes(x=Date,y= mean_land_value, color = region))+
  geom_line()

```

**Exercise 3. Life expectancy and GDP per capita 1952-2007**

This exercise uses the `gapminder` dataset from the `gapminder` package. It describes the life expectancy `(lifeExp`), GDP per capita ('gdpPercap`), and population (`pop`) of 142 countries from 1952 to 2007. These countries can be grouped into 5 continents. As a reminder, **reproduce the following plots exactly as shown**.

```{r}
kable(head(gapminder))
```

 
```{r}
Scatter_pl<-gapminder %>% 
ggplot(mapping =aes(x=gdpPercap, y = lifeExp))+
  geom_point()
Scatter_pl
```
 
 
 
**3.2 Add a smoothing line to the previous plot.**

```{r,echo=FALSE, message = FALSE}
Scatter_pl + 
geom_smooth()
```

 
**3.3 Show each continent in a different color, and fit a separate smoothing line to each continent to identify differences in this relationship between continents. Turn off the confidence intervals.**


```{r,echo = FALSE, message=FALSE}
Scatter_pl + 
 geom_point(aes(color = continent))+
  geom_smooth(aes(color = continent), se = FALSE) 

```
 
 
 
**Use faceting to solve the problem. Show the confidence intervals in this plot**

```{r, echo FALSE, message = FALSE}
Scatter_pl + 
 geom_point(aes(color = continent))+
  geom_smooth(aes(color = continent))+
  facet_wrap(~continent)
```

 
 
 

 
 
 
 
 