---
title: "Assignment 4: Data transformation with dplyr and visualization with ggplot "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Instructions: Please read through this before you begin**

* This assignment is due by **10pm on Monday 10/19/20**.

* For this assignment, please **reproduce this markdown file** using R markdown. This includes the followings:

  + **Reproduce this markdown template.** Pay attention to all the formatin gin this file, including bullet points,       bolded characters, inserted code chunks, headings, text colors, blank lines, etc.

  + **Transform the data as instructed.** Try to use `tidyverse` functions even if you are more comfortable with         base-R solutions.Show **the first 6** lines of the transformed data in a table through RMarkdown **using the         kable() function**, as shown in this markdown file.
  
  + **Reproduce the plots exactly as shown in this html file.** In two cases where the plot is not shown (Excercises     3.7 and 3.9), generate plots that you think can best answer the question.
  + Have all your code embedded within the R markdown file, and show **BOTH yourcode and plots** in the knitted           markdown file
  
  + When a verbal response is needed, answer by editing the part in the R markdwon template where it says "Write your    response here".
  
  + Use R Markdwon functionalities to **hide messages and warnings when needed**. (Suggestion: messages and warnings    can often be informative and important, so please examine them carefully and only turn them off when you finish     the exercise).
  
* Please name your R markdown file `assignment_4.Rmd` and the knitted markdown file `assignmrent_4.md`. Plese push     both files to your class GitHub repository.
  
First, load all the required packages with the following code. Install them if they are not installed yet.
  
```{r, message=FALSE}
library(tidyverse)
library(knitr)
library(gapminder) #
```

  
**Exercise 1. Theopylline experiment** 
  
This excercise uses the `Theoph` data frame (comes with your R installation), which has 132 rows and 5 columns of data from an experiment on the pharmacokinetics of the anti-asthmatic drug theophylline. Tweleve subjects were given oral doses of theophylline then serum concentrations were measured at 11 time points over the next 25 hours. You can learn more about this dataset by running `?Theoph` 

Have a look at the data structure

```{r}
kable(head(Theoph))
```


**1.1 Select columns that contain a lower case"t" in the `Theoph` dataset. Do not mannually list all the columns to include.**


```{r}

kable(head(Theoph %>% 
  select(Subject, Wt)))


```

```{r}
kable(head(select(Theoph, contains(("t"))))) # this selects the upper case T why?
```



**Rename the `Wt` column to `Weight` and `conc` column to `Concentration` in the `Theoph` dataset.**

```{r}
kable(head(Theoph %>% 
select(everything(), Weight = Wt, Concentration =conc)))

```





  
  
  
  
  
  
  
  